Unifying intraspecific and interspecific variation in tropical tree mortality

### 2017-06-08 ms/ms.Rnw
Diverse minor edits throughout
- adding units and details to axis labels
- tightening paragraphs to save a line, where possible
- more details where needed

\author[a,b,1]{James S Camac}
\author[c]{Richard Condit}
\author[d]{Richard G FitzJohn}
\author[e]{Lachlan McCalman}
\author[e]{Daniel Steinberg}
\author[a]{Mark Westoby}
\author[f]{S Joseph Wright}
\author[a,g]{Daniel S Falster}

@@

% Please add here a significance statement to explain the relevance of your work
% 120 words
\significancestatement{Tree mortality is a fundamental demographic process affecting forest dynamics and carbon cycling. Here, for the first time, we use observed survival records over a 15 year period from more than 180,000 individuals, to simultaneously estimate growth-dependent and growth-independent (i.e. stochastic) mortality across 203 tropical forest tree species. We found that growth-dependent mortality was the predominate factor influencing tree mortality rates at Barro Colorado Island. Furthermore, we found that wood density most strongly influenced mortality rates by decreasing growth-independent mortality. However, wood density only accounted for a small fraction of the overall species variability observed in growth-independent mortality rates, suggesting that there must be other species traits that strongly affect mortality. Which traits and how many of them, however, remains unknown.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{D.F. conceived idea; all authors designed research; J.C. \& R.F. performed analysis; R.C. \& S.J.W. collected data. J.C \& D.F. wrote the manuscript}

@@

Data deposition: Code and data reproducing the results in this paper are available at \href{https://github.com/traitecoevo/mortality_bci}{github.com/traitecoevo/mortality\_bci}.
}

@@

Two difficulties arise when studying tree mortality in tropical rainforests. The first is that large population sizes and long periods of observation are required to make inferences into how various factors affect mortality rates\cite{Poorter:2008iu}. This requirement arises in part because the observable outcome -- alive vs. dead -- is one-step removed from the variable we ideally want to measure: $\lambda_i(t)$, the instantaneous rate of mortality for individual $i$ at time $t$, also called the ``hazard function''. In classic survival analysis\cite{Kleinbaum:2012tn}, the probability an individual dies between times $t_2$ and $t_3$ is a function of $\lambda_i(t)$:
\begin{equation} \label{eq:mortality_model}
p_{i,t_2\rightarrow t_3} = 1 - \exp\left(- \int_{t_2}^{t_3} \lambda_i(t) {\rm d}t\right).
\end{equation}
Here $\int_{t_2}^{t_3} \lambda_i(t) {\rm d}t$ is the ``cumulative hazard'' between $t_2$ and $t_3$\cite{Kleinbaum:2012tn}. The observed survival outcome $S_{i, t_2\rightarrow t_3}$ (0 = alive, 1 = died) is then a realisation of this probability
\begin{equation}
S_{i,t_2\rightarrow t_3} \sim \textrm{Bernoulli}(p_{i,t_2\rightarrow t_3}).
\end{equation}
Since trees are long-lived and we are trying to estimate the shape of a continuous hazard function from binary data, large sample-sizes are required. As such, detailed studies of tree mortality have only recently become possible, with the accumulation of growth and survival data from repeat surveys spanning several decades in plots containing thousands of individuals\cite{Anderson-Teixeira-2015a}.

@@

In this study, we attempt to reconcile intra-specific and inter-specific factors as well as partition instantaneous mortality rates into growth-dependent and growth-independent rates. We achieve this by evaluating the following, unified hazard function, incorporating individual-, species-, and census-level effects:
\begin{equation} \label{eq:model}
\lambda_{i}(t) = \bigg(\overbrace{\gamma_{s[i]}}^{\text{Growth independent}} + \overbrace{\alpha_{s[i]} \, \textrm{e}^{-\beta_{s[i]} \, X_i(t)}}^{\text{Growth dependent}}\bigg) \times \overbrace{\delta_{t}}^{\text{Census}}.
\end{equation}
Eq. \ref{eq:model} allows for additive growth-independent and growth-dependent effects, includes a negative exponential effect of growth rate, and allows for mortality to vary temporally among censuses, via the random effect $\delta_{t}$. Further, the parameters $\alpha_s$, $\beta_s$ and $\gamma_s$ are allowed to vary by species. Here we include an effect of a species trait (wood density, $\rho$), as well as a species random effect that captures any additional species-level differences not represented $\rho$ (for details see Methods).

To validate this model, we fit a series of models with increasing complexity, with eq. \ref{eq:model} being the most complex, and compare their skill in predicting patterns of tropical tree mortality for \Sexpr{n_inds(BCI_model_dataset_true_unique)} individuals from \Sexpr{n_spp(BCI_model_dataset_true_unique)} tropical rainforest species at Barro Colorado Island (BCI), Panama (Fig. \ref{fig1}). The data are repeat censuses of stem diameter and tree status (alive vs. dead) taken over a 15 yr period (Fig. \ref{fig1}A). We compare the skill of different hazard functions in predicting outcomes in novel data (i.e. not used in model calibration) via 10-fold cross-validation (Fig. \ref{fig1}D) \cite{Hooten:2015, Stone:1997tg}. Evaluating models in this way is computationally expensive, making it impossible to run all possible model formulations. We therefore fit models across five iterative stages of model development. At each stage, the best model from the previous stage was taken as input for the next stage.

In stage 1 we ask whether mortality rates vary substantially between censuses to establish whether a census effect is required. The null model is simply a constant, invariant with respect to species, growth rate, or year.
In stage 2 we assess whether a hazard function including both growth-dependent and independent terms outperform a function including only one of these? Species effects were excluded, so we are simply asking which of three functional forms for $\lambda_i(t)$ (Fig. \ref{fig1}B) best represents the data. The simplest form (Fig. \ref{fig1}B, left) assumes a constant growth-independent hazard rate. The second form (Fig. \ref{fig1}B, middle) assumes the risk of dying eventually declines towards an asymptote of zero as growth increases. The third form (eq. Fig. \ref{fig1}B, right) is the summation of the two previous models, which allows $\lambda_i(t)$ to decrease exponentially with increasing growth rate, but, unlike a standard negative exponential, asymptoting at some baseline hazard $\textgreater 0$. Combined, these three models capture a variety of functional responses previously proposed, including effects represented in current vegetation models \cite{Moorcroft:2001ws,McDowell:2011dr}, which have not previously been systematically compared. We also investigate which growth measure (stem diameter or stem area increment) more skilfully predicts growth-dependent mortality.
In stage 3, we examined whether including wood density (a species-level trait) improved model skill, and if so, whether the effect of wood density was primarily on growth-dependent or growth-independent hazards, or both? Finally, we fit a model which allowed parameters to vary by species-level variation that was otherwise not captured by wood density. This allowed us to ask what proportion of inter-specific hazard variability is explained by wood density. Using the final `best' model we also conducted some posthoc tests, to determine how species-level parameters were associated with their maximum size and light requirement.

@@

\begin{figure}[!h]
\includegraphics{figures/fig2.pdf}
\caption{Mean ($\pm$ 95\% credible intervals) logarithmic loss whereby lower values equal higher predictive skill.
A) Five sequential stages of model selection with each stage increasing in complexity. Null: null model ($\gamma$); Census: inclusion of census effects ($\delta_t$); Growth rate: the inclusion of growth rate. This includes two possible hazard functions: growth-dependent only hazard ($(\alpha e^{-\beta X_{i}})\delta_t$) and baseline + growth-dependent hazard with census effects ($(\alpha e^{-\beta X_{i}} + \gamma)\delta_t$); WD: Inclusion of wood density effect ($\rho$) on $\gamma$; Species: the inclusion of species random effects $s$ on all parameters.
B) Change in logloss under all wood density\textendash parameter combinations. Colors represent growth measure used: white = no growth measure, grey = basal area growth, black = dbh growth. Symbols represent general functional form: null/baseline (circle), growth\textendash dependent hazard model (triangle) and model incorporating both baseline and growth\textendash dependent hazards (square). Note: errors are too small to see in panel A.}
\label{fig2}
\end{figure}

\subsection*{Growth hazard functions}
Comparing the three hazard functions in Fig. \ref{fig1}B, we found that the third function -- with both growth-dependent and growth-independent terms, i.e. $(\alpha e^{-\beta X_{i}} + \gamma)\delta_t$ -- significantly outperformed both the baseline only or growth-dependent only functions (Fig. \ref{fig2}A). Moreover, we found that predictive skill was higher when using stem diameter growth over stem-area growth (Fig. \ref{fig2}A).

\begin{figure}[h!]
\includegraphics{figures/fig3.pdf}
\caption{Relationship between species baseline mortality rate and wood density. Points are estimated mean baseline mortality rates $\pm$ 95\% credible intervals for each of the 203 BCI species used in this study. Blue trendline with grey shading shows average ($\pm$ 95\% credible intervals) expected relationship.}
\label{fig3}
\end{figure}

\subsection*{Effect of wood density on mortality rates}
Including wood density ($\rho$) as an effect on either baseline or growth-dependent parameters significantly improved model skill relative to a hazard function without such effects (Fig. \ref{fig2} A,B). The most parsimonious model, with highest predictive skill, was that which attributed the wood density effect to the baseline hazard term ($\gamma$; Fig. \ref{fig2} B). Wood density was therefore negatively correlated with a species' baseline mortality rate (Fig. \ref{fig3}).

@@

\subsection*{Proportion of variation explained}
By partitioning the modelled variation between our two random effects (i.e. species and census; Fig. \ref{fig5}A), we found that approximately \Sexpr{extract_prop_explained(param_variance_explained, I('species'))}\% of variation in predicted mortality rates was due to species differences, only \Sexpr{extract_prop_explained(param_variance_explained, I('wood_density'))}\% of which was captured by wood density. By contrast, census effects accounted for approximately \Sexpr{extract_prop_explained(param_variance_explained, I('census'))}\% of the total predicted variation in mortality rates. When we partitioned modelled variation into the two independent hazards we found that the growth-dependent hazard (i.e. the part of the model allowed to vary by individual growth) accounted for \Sexpr{extract_prop_explained(param_variance_explained, I('growth_dependent'))}\% of the overall variation in mortality rates, and the remainder was attributed to the baseline hazard.

\subsection*{Mortality rates, maximum species size and species light requirements}
We found that both $\alpha$, the effect of low growth rate, and $\gamma$, the baseline hazard term, were weakly and positively correlated with gap index (our measure of species light requirement; Fig. \ref{fig6} A,C, E). By contrast, $\beta$ (the parameter that defines the exponential decay with increasing growth rate) was negatively correlated gap index. The combination of these correlations suggest that species which predominately recruit in gaps (i.e. higher gap index) were more prone to death due to both low growth and stochastic chance. They also required faster growth rates to achieve a given mortality rate relative to a species that recruits in non-gaps. We detected no strong correlations between estimated species parameters and their associated maximum DBH (Fig. \ref{fig6} B,D,F,H).

\begin{figure}[t]
\includegraphics{figures/fig5.pdf}
\caption{Proportion of predicted variation explained.
A) Predicted variation explained by wood density (grey) as well as species and census random effects.
B) Predicted variation explained by growth\textendash dependent and baseline (stochastic) hazards.}
\label{fig5}
\end{figure}

@@

The findings of this study provide empirical support for dynamic vegetation models that estimate mortality as the sum of growth-dependent and growth-independent hazards\cite{Pacala:1996vm, Moorcroft:2001ws, McDowell:2011dr}. We show that regardless of growth measure, incorporating both hazards significantly improves model predictive skill. This is because the additive effects of both hazards explicitly allow models to account for two independent causes of plant mortality. The growth-dependent hazard allows for deaths associated with low carbon budgets, and as a consequence, incorporates intra-specific variability attributed to carbon related stresses (e.g. competition, parasites, herbivory). By contrast, the growth-independent hazard accounts for deaths caused by events that arise irrespective of an individuals growth rate (e.g. windthrow).

@@

How the length between censuses affect the estimation of growth-dependent and growth-independent hazards should also be examined. Large census intervals are likely to lead to underestimation of growth-dependent mortality, and consequently, overestimate the relative contribution of growth-independent hazards. This is because if a census interval is too wide, it will fail to account for deaths attributed to rapid declines in growth caused by senescence (especially for short-lived plants) or events such as drought. Because our analysis estimated an individual's past growth from 5 year census intervals, and because it included short-lived species (i.e. those with high annual mortality rates; see Fig. \ref{fig4}) and a drought period between 1995 and 2000, our estimates of growth-dependent and growth-independent hazards are likely to be biased.

Lastly, research should also examine how different growth measures influence predictive skill of mortality models. We have no clear explanation as to why DBH growth outperformed basal area growth in terms of model predictive skill, and we are unsure how other common measures such as biomass growth\cite{Chave:2005hz} will compare.

@@

\subsection*{Posthoc correlations}
We calculated a gap index as a measure of a species' light dependence, using annual canopy census data collected during 1985\textendash1990 and 1990\textendash1995. The canopy census recorded the presence of vegetation in six height intervals: 0\textendash2, 2\textendash5, 5\textendash10, 10\textendash20, 20\textendash30 and > 30 m, in all 5x5 m subplots across the 50 ha plot. For each subplot, we calculated the minimum number of strata > 2 m containing vegetation; and then transformed this to a light index ranging from 0 (dense shade) to 1 (gap). We then calculated a weighted sum for each subplot, using a 3 x 3 cell moving window, where the central cell was assigned a weight of 8, and the neighbouring cells were assigned weights of 1. The gap index for each species was then calculated by averaging the light level encountered by new saplings as they appeared in the census (Fig. S2).

@@

\acknow{
J.C was supported by the Science and Industry Endowment Fund (SIEF; RP04-174). D.F. and M.W were supported by fellowships from the Australian Research Council (ARC). We thank Rich FitzJohn, Lachlan McCalman and Daniel Steinberg for computational advice.  Fig. 1a uses images by Tracey Saxby, IAN Image Library.}

>>>>

\author[a,b,1]{James S Camac}
\author[c]{Richard Condit}
\author[d]{Richard G FitzJohn}
\author[e]{Lachlan McCalman}
\author[e]{Daniel Steinberg}
\author[a]{Mark Westoby}
\author[f]{S Joseph Wright}
\author[g]{Daniel S Falster}

@@

% Please add here a significance statement to explain the relevance of your work
% 120 words
\significancestatement{Tree mortality is a fundamental demographic process affecting forest dynamics and carbon cycling. Here, for the first time, we use observed survival records over a 15 year period from more than 180,000 individuals, to simultaneously estimate growth-dependent and growth-independent (i.e. stochastic) mortality across 203 tropical forest tree species. We found that growth-dependent mortality was the predominate factor influencing tree mortality rates at Barro Colorado Island. Furthermore, we found that wood density most strongly influenced mortality rates by decreasing growth-independent mortality. However, wood density only accounted for a small fraction of the overall species variability observed in growth-independent mortality rates, suggesting that there must be other species traits that strongly affect mortality. Which traits remains unknown.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{D.F. conceived idea; all authors designed research; J.C. \& R.F. performed analysis; R.C. \& S.W. collected data. J.C \& D.F. wrote the manuscript}

@@

Data deposition: Code reproducing the results is available at \href{https://github.com/traitecoevo/mortality_bci}{github.com/traitecoevo/mortality\_bci}. Data from BCI is available at DOIs \href{http://doi.org/10.5479/data.bci.20130603.}{10.5479/data.bci.20130603.}, \href{http://doi.org/10.5479/data.bci20140711}{10.5479/data.bci20140711}.
}

@@

Two difficulties arise when studying tree mortality in tropical rainforests. The first is that large population sizes and long periods of observation are required to make inferences into how various factors affect mortality rates\cite{Poorter:2008iu}. This requirement arises in part because the observable outcome -- alive vs. dead -- is one-step removed from the variable we ideally want to measure: $\lambda_i(t)$, the instantaneous rate of mortality for individual $i$ at time $t$, also called the ``hazard function''. In classic survival analysis\cite{Kleinbaum:2012tn}, the probability an individual dies between times $t_2$ and $t_3$ is a function of $\lambda_i(t)$:
\begin{equation} \label{eq:mortality_model}
p_{i,t_2\rightarrow t_3} = 1 - \exp\left(- \int_{t_2}^{t_3} \lambda_i(t) {\rm d}t\right).
\end{equation}
Here $\int_{t_2}^{t_3} \lambda_i(t) {\rm d}t$ is the ``cumulative hazard'' between $t_2$ and $t_3$\cite{Kleinbaum:2012tn}. The observed survival outcome $S_{i, t_2\rightarrow t_3}$ (0 = alive, 1 = died) is then a realisation of this probability
\begin{equation}
S_{i,t_2\rightarrow t_3} \sim \textrm{Bernoulli}(p_{i,t_2\rightarrow t_3}).
\end{equation}
As trees are long-lived and we are trying to estimate the shape of a continuous hazard function from binary data, large sample-sizes are required. Detailed studies of tree mortality have thus only recently become possible, with the accumulation of growth and survival data from repeat surveys spanning several decades in plots containing thousands of individuals\cite{Anderson-Teixeira-2015a}.

@@

Here, we attempt to reconcile intra- and inter-specific factors as well as partition instantaneous mortality rates into growth-dependent and growth-independent rates. We achieve this by evaluating the following, unified hazard function, incorporating individual-, species-, and census-level effects:
\begin{equation} \label{eq:model}
\lambda_{i}(t) = \bigg(\overbrace{\gamma_{s[i]}}^{\text{Growth independent}} + \overbrace{\alpha_{s[i]} \, \textrm{e}^{-\beta_{s[i]} \, X_i(t)}}^{\text{Growth dependent}}\bigg) \times \overbrace{\delta_{t}}^{\text{Census}}.
\end{equation}
Eq. \ref{eq:model} allows for additive growth-independent and growth-dependent effects, includes a negative exponential effect of growth rate, and allows for mortality to vary temporally among censuses, via the random effect $\delta_{t}$. Further, the parameters $\alpha_s$, $\beta_s$ and $\gamma_s$ are allowed to vary by species. Here we include an effect of a species trait (wood density, $\rho$), as well as a species random effect that captures any additional species-level differences not represented $\rho$ (for details see Methods).

To validate this model, we fit a series of models with increasing complexity (eq. \ref{eq:model} being the most complex) and compare their skill in predicting patterns of tropical tree mortality for \Sexpr{n_inds(BCI_model_dataset_true_unique)} individuals from \Sexpr{n_spp(BCI_model_dataset_true_unique)} tree species at Barro Colorado Island (BCI), Panama (Fig. \ref{fig1}). The data are repeat censuses of stem diameter and tree status (alive vs. dead) taken over a 15 yr period (Fig. \ref{fig1}A). We compare the skill of different hazard functions in predicting outcomes in novel data (i.e. not used in model calibration) via 10-fold cross-validation (Fig. \ref{fig1}D) \cite{Hooten:2015, Stone:1997tg}. Evaluating models in this way is computationally expensive, making it impossible to run all possible model formulations. We therefore fit models across five stages of model development. At each stage, the best model from the previous stage was taken as input for the next stage.

In stage 1 we ask whether mortality rates vary substantially between censuses, to establish whether a census effect is required. The null model is simply a constant, invariant with respect to species, growth rate, or year.
In stage 2 we ask whether a hazard function including both growth-dependent and independent terms outperforms a function including only one of these? Species effects were excluded, so we are simply asking which of three functional forms for $\lambda_i(t)$ (Fig. \ref{fig1}B) best represents the data. The simplest form (Fig. \ref{fig1}B, left) assumes a constant (growth-independent) hazard rate. The second form (Fig. \ref{fig1}B, middle) assumes the risk of dying declines towards an asymptote of zero as growth increases. The third form (eq. Fig. \ref{fig1}B, right) is the summation of the two previous models; this allows $\lambda_i(t)$ to decrease exponentially with increasing growth rate, but, unlike B, asymptoting at some baseline rate $\textgreater 0$. Combined, these three models capture a variety of functional responses previously proposed, including effects represented in current vegetation models \cite{Moorcroft:2001ws,McDowell:2011dr}, which have not previously been systematically compared. We also investigate which growth measure (stem diameter or stem area increment) more skilfully predicts growth-dependent mortality.
In stage 3, we examined whether including wood density (a species-level trait) improved model skill, and if so, whether the effect of wood density was on growth-dependent or growth-independent hazard, or both? Finally, we fit a model that allowed parameters to vary by species-level effects not captured by wood density. This allowed us to ask what proportion of inter-specific hazard variability is explained by wood density. Using the final `best' model we also conducted some posthoc tests, to determine how species-level parameters were associated with their maximum size and light requirement.

@@

\begin{figure}[!h]
\includegraphics{figures/fig2.pdf}
\caption{Comparing predictive skill of alternative hazard functions. Values are mean ($\pm$ 95\% credible intervals) logarithmic loss, with lower values implying greater skill.
A) Five sequential stages of model selection with each stage increasing in complexity. Null: constant ($\gamma$); Census: inclusion of census effects ($\delta_t$); Growth rate: the inclusion of growth rate. This stage compares two hazard functions: growth-dependent only hazard ($(\alpha e^{-\beta X_{i}})\delta_t$) and growth-independent + growth-dependent hazard ($(\alpha e^{-\beta X_{i}} + \gamma)\delta_t$); WD: Inclusion of wood density effect ($\rho$) on $\gamma$; Species: the inclusion of species random effects $s$ on all parameters.
B) Comparing predictive skill for wood density parameter combinations. Colors represent growth measure used: white = no growth measure, grey = basal area growth, black = dbh growth. Symbols represent functional form: null and growth-independent (circle), growth-dependent hazard (triangle), both growth-independent and growth-dependent hazards (square). Errors in A are too small to see.}
\label{fig2}
\end{figure}

\subsection*{Growth hazard functions}
Comparing the three hazard functions in Fig. \ref{fig1}B, we found that the third function -- with both growth-dependent and growth-independent terms, i.e. $(\alpha e^{-\beta X_{i}} + \gamma)\delta_t$ -- significantly outperformed both the growth-independent only or growth-dependent only functions (Fig. \ref{fig2}A). Moreover, we found that predictive skill was higher when using stem diameter growth over stem-area growth (Fig. \ref{fig2}A).

\begin{figure}[h!]
\includegraphics{figures/fig3.pdf}
\caption{Relationship between species growth-independent mortality rate and wood density. Points are estimated mean baseline mortality rates $\pm$ 95\% credible intervals for each of the 203 BCI species used in this study. Blue trendline with grey shading shows average ($\pm$ 95\% credible intervals) expected relationship.}
\label{fig3}
\end{figure}

\subsection*{Effect of wood density on mortality rates}
Including wood density as an effect on either growth-independent or growth-dependent parameters significantly improved model skill, relative to a model without such effects (Fig. \ref{fig2} A,B). The most parsimonious model, with highest predictive skill, was that attributing the wood density effect to the growth-independent hazard term ($\gamma$; Fig. \ref{fig2} B). Thus wood density was correlated with a species' growth-independent mortality rate (Fig. \ref{fig3}).

@@

\subsection*{Proportion of variation explained}
By partitioning the modelled variation between our two random effects (i.e. species and census; Fig. \ref{fig5}A), we found that approximately \Sexpr{extract_prop_explained(param_variance_explained, I('species'))}\% of variation in predicted mortality rates was due to species differences, only \Sexpr{extract_prop_explained(param_variance_explained, I('wood_density'))}\% of which was captured by wood density. By contrast, census effects accounted for approximately \Sexpr{extract_prop_explained(param_variance_explained, I('census'))}\% of the total predicted variation in mortality rates. When we partitioned modelled variation into the two independent hazards we found that the growth-dependent hazard (i.e. the part of the model allowed to vary by individual growth) accounted for \Sexpr{extract_prop_explained(param_variance_explained, I('growth_dependent'))}\% of the overall variation in mortality rates, and the remainder was attributed to the growth-independent hazard.

\subsection*{Mortality rates, maximum species size and species light requirements}
We found that $\alpha$, the effect of low growth rate, and $\gamma$, the growth-independent hazard term, were weakly and positively correlated with gap index, a measure of species light requirement (Fig. \ref{fig6} A,C, E). By contrast, $\beta$ (the exponential decay with increasing growth rate) was negatively correlated gap index. The combination of these correlations suggest that species which predominately recruit in gaps (i.e. higher gap index) were more prone to death due to both low growth and stochastic chance. They also required faster growth rates to achieve a given mortality rate relative to a species that recruits in non-gaps. We detected no strong correlations between estimated species parameters and their associated maximum DBH (Fig. \ref{fig6} B,D,F,H).

\begin{figure}[t]
\includegraphics{figures/fig5.pdf}
\caption{Proportion of predicted variation in 5-yr survival explained by different covariates.
A) Wood density (grey) as well as total species and census random effects.
B) Growth-dependent and growth-independent (baseline) hazards.}
\label{fig5}
\end{figure}

@@

The findings of this study provide empirical support for dynamic vegetation models that estimate mortality as the sum of growth-dependent and growth-independent hazards\cite{Pacala:1996vm, Moorcroft:2001ws, McDowell:2011dr}. We show that regardless of growth measure, incorporating both hazards significantly improves model predictive skill. This is because the additive effects of both hazards explicitly allow models to account for two independent causes of plant mortality. The growth-dependent hazard allows for deaths associated with low carbon budgets, and as a consequence, incorporates intra-specific variability attributed to carbon related stresses (e.g. competition, parasites, herbivory). By contrast, the growth-independent hazard accounts for deaths caused by events that arise irrespective of an individual's growth rate (e.g. windthrow, lightning strike).

@@

Whether the length between censuses biases estimation of growth-dependent and growth-independent hazards should also be examined. Long census intervals may lead to underestimation of growth-dependent mortality, and consequently, overestimate the relative contribution of growth-independent hazards. This is because if a census interval is too wide, it will fail to account for deaths attributed to rapid declines in growth caused (especially for short-lived plants) or events such as drought. Because our analysis estimated an individual's past growth from 5 year census intervals, and because it included short-lived species (i.e. those with high annual mortality rates; see Fig. \ref{fig4}) and a drought period between 1995 and 2000, our estimates of growth-dependent and growth-independent hazards are likely to be biased.

Lastly, research should also examine how different growth measures alter the predictive skill of mortality models. We found that DBH growth outperformed basal area growth, but have as yet  no clear explanation as to why, or how other common measures such as biomass growth\cite{Chave:2005hz} will compare.

@@

\subsection*{Posthoc correlations}
We calculated a gap index as a measure of a species' light dependence, using canopy census data collected from 1985\textendash1995\cite{Hubbell:2014}. The canopy census recorded, in all 5x5 m subplots across the 50 ha plot, the presence of leaf in six height intervals between 0, 2, 5, 10, 20, 30 and > 30 m. For each subplot, we calculated the minimum number of strata > 2 m containing leaf; and then transformed this to a light index ranging from 0 (dense shade) to 1 (gap). We then calculated a weighted mean for each subplot, averaging each cell's value with mean of the 8 neighbouring cells. The gap index for each species is then the average light level encountered by new saplings appearing in the census (Fig. S2).

@@

\acknow{We thank H. Muller-Landau and XX anonymous reviewers for feedback and B. Carpenter for technical advice.  We acknowledge S. Hubbell, R. Foster, R. Pérez, S. Aguilar, S. Lao, S. Dolins, \& hundreds of field workers for their contribution; and the National Science Foundation, Smithsonian Tropical Research Institute, \& MacArthur Foundation for funding the design, collection, quality control and management of long-term growth data at BCI. Fig. 1a uses images by Tracey Saxby, IAN Image Library. J.C., R.F., L.M. and D.S. were supported by the Science and Industry Endowment Fund (SIEF; RP04-174). D.F. and M.W. were supported by fellowships from the Australian Research Council.}

### 2017-06-08 ms/supp.Rnw
Diverse minor edits throughout
- adding units and details to axis labels
- tightening paragraphs to save a line, where possible
- more details where needed

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1]{James S Camac}
\author[a]{Mark Westoby}
\author[b]{S Joseph Wright}
\author[c]{Daniel S Falster}

\affil[a]{Department of Biological Sciences, Macquarie University NSW 2109, Australia}
\affil[b]{Smithsonian Tropical Research Institute, Apartado 0843–03092, Balboa, Panama}
\affil[c]{Evolution and Ecology Research Centre, School of Biological, Earth and Environmental Science, University of New South Wales}

>>>>

\author[a,1]{James S Camac}
\author[b]{Richard Condit}
\author[c]{Richard G FitzJohn}
\author[d]{Lachlan McCalman}
\author[d]{Daniel Steinberg}
\author[a]{Mark Westoby}
\author[e]{S Joseph Wright}
\author[f]{Daniel S Falster}

\affil[a]{Department of Biological Sciences, Macquarie University NSW 2109, Australia}
\affil[b]{Field Museum of Natural History and Morton Arboretum, Chicago, Illinois, USA}
\affil[c]{Imperial College, London, United Kingdom}
\affil[d]{Data61, CSIRO, Australia}
\affil[e]{Smithsonian Tropical Research Institute, Apartado 0843–03092, Balboa, Panama}
\affil[f]{Evolution and Ecology Research Centre, University of New South Wales, Australia}
